Inspiring Confidence and Easing Insecurity in the Pre-election period: The Effect of Recruiting Military Veterans on Confidence in 2024 US Elections

### 2025-07-07 docs/index.qmd
revisions

# custom function to return a certain selection of logistic regression model
# goodness-of-fit statistics
lrm.gof.stats <- function(x, ...){

  out <- data.frame(
    null.deviance = x$null.deviance,
    df.null = x$df.null,
    logLik = as.numeric(stats::logLik(x)),
    AIC = stats::AIC(x),
    BIC = stats::BIC(x),
    deviance = stats::deviance(x),
    df.residual = stats::df.residual(x),
    nobs = stats::nobs(x),
    chisq = as.numeric(x$null.deviance - stats::deviance(x)),
    df = as.numeric(x$df.null - stats::df.residual(x)),
    'P(>chi)' = pchisq(q=x$null.deviance - stats::deviance(x),
                       df = x$df.null - stats::df.residual(x), 
                       lower.tail = F)
  )
  out <- dplyr::as_tibble(out)
  return(out)
}

@@

Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by prior impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve trust and confidence in election administration, and lessen expectations of electoral fraud. The next section elaborates on why military veterans are of particular interest in this regard.

## Why Veterans

Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Especially since staffing issues have been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

The reason why veterans arose as a special subset of the population to consider for this inquiry is because there was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [-@cooter2024; see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

That being said, the general public perception, attitudes, or even stereotypes about military veterans are significant to consider. Although military recruitment shows a downward trend as of late, public perceptions of veterans are overwhelmingly positive [@kleykamp2023]. Veterans are a particularly potent group where mention of veteran status seems to have a positive, calming, or nullifying effect on attitudes. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to a man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015].

## Theory and Hypotheses

Considering the generally positive perception afforded to veterans by the public, it is reasonable to suspect that efforts that promote recruitment of veterans and their family members to work as election staff and volunteers would boost public *confidence in elections administration*. Indeed, that is the primary hypothesis of this study. Although citizens may vote in various ways across the country (e.g., by mail, ballot drop box, in-person), the simple announcement that election officials are engaging in efforts to recruit veterans to work as election staff may boost confidence in elections administration regardless of how, or whether, an individual plans to vote. Formally,

@@

@@

```{r}
#| label: fig-coef1
#| layout-nrow: 2
#| fig-height: 3.5
#| fig-pos: 'H'
#| fig-cap: "Confidence in Elections by Experiment Condition"
#| fig-subcap: 
#|   - "Confidence by Treatment Condition"
#|   - "Trust and Distrust by Treatment Condition"

@@

@@

conf.coef1
conf.coef2

@@

@fig-coef1 displays standard difference estimates of confidence in elections by treatment condition (control condition as reference). @fig-coef1-1 displays two models for confidence in elections in Maricopa County, AZ, and one's local area by treatment condition; @fig-coef1-2 displays the same except *confidence* is decomposed into its two components, *trust* and *distrust*. These figures merely illustrate the effect of the treatment with respect to the relationship between trust and distrust; ultimately confidence improves.

@@

::: {.content-visible when-format="bookup-html" unless-format="pdf"}

```{r}
#| tbl-cap: "Confidence in Elections by Treatment and Legitimacy Beliefs"
#| tbl-pos: H

@@

:::

::: {.content-visible when-format="pdf" unless-format="bookup-html"}

```{r}
#| tbl-cap: "Confidence in Elections by Treatment and Legitimacy Beliefs"
#| tbl-pos: H
#| label: tbl-legit
#| results: asis

texreg::texreg(
  list(m1, m2),
  single.row = T,
  caption = NULL,
  caption.above = T,
  use.packages = F,
  table = F, # set `table = FALSE`. Quarto handles table cross-ref env
  custom.model.names = c("Confidence in AZ Elections", "w/Interaction"),
  custom.coef.names = c(
    "(Intercept)",
    "Treatment",
    "Not Legitimate",
    "Republican",
    "Democrat",
    "Not Legitimate * Treatment"
    )
)

:::

The first model in Table \ref{tbl-legit} shows that overall confidence in elections among those who read the treatment vignette was significantly (statistically) greater than those who read the control, on average and netting out the effects of partisanship and legitimacy beliefs. Democratic partisanship remains marginally significant at $90\%$ confidence level ($p = 0.055$). Not believing the 2020 election of Joe Biden was legitimate is negatively associated with confidence in elections. In order to assess the magnitude of this interaction effect of the treatment on those who refuted the 2020 election results, I ran a model where confidence in elections was regressed on the interaction between the treatment and legitimacy beliefs, controlling for partisanship.

@@

Thus far, publicized announcements that election officials will be targeting veterans in their recruitment efforts does quite a bit to improve confidence in elections outside of one's local area (or, at least for elections held in Maricopa County, AZ), especially among those who believed the 2020 election results were illegitimate. While distrust is one concern during the pre-election period, it is not the only concern that arises in anticipation of US elections; the public must feel confident that voters are safe to cast a ballot in person. The survey included questions^[See @tbl-4 in Appendix A] that allows me to assess whether those in the treatment condition expressed less concern for violence and more confidence in voter safety compared to those in the control condition. 
```{r}
#| label: fig-safety-bars
#| layout-ncol: 1
#| layout-nrow: 2
#| fig-cap: "Safety Concerns and Confidence by Treatment Condition"
#| fig-subcap: 
#|   - "Concern for Violence while voting in Maricopa County, AZ Elections"
#|   - "Confidence in Voter Saftey at Election Sites in Maricopa County, AZ"

@@

@@

Briefly reviewing @fig-safety-bars, those in the treatment condition felt less concern over the potential for violence compared to those in the control. As for confidence in voter safety at election sites, it appears that more respondents in the treatment condition expressed that they were "Very confident" compared to those in the control, by approximately $8$ percentage points.

@@

Those in the treatment condition had lower probability to express that they were 'Very' or 'Somewhat' concerned for violence in Maricopa County, AZ Elections, yet higher probability of expressing 'Not too' or 'Not at all' concerned. As for confidence in voter safety, moving from treatment to control, the difference in predicted probability of a respondent selecting "Very Confident" was $0.08$. @tbl-preds-diff1 and @tbl-preds-diff2 show the significant differences in predicted probabilities at 95$\%$ confidence^[The same models were run for items pertaining to safety concerns in one's local area but results of the treatment were not significantly different from the control. See Appendix D for full results table of ordinal logistic regression].

@@

```{r}
#| label: tbl-preds-diff1
#| tbl-colwidths: 80
#| tbl-cap: "Difference in Predicted Probabilities in Concern for violence, Threats of Violence, or Intimidation between Treatment and Control"

@@

```{r}
#| label: tbl-preds-diff2
#| tbl-colwidths: 80
#| tbl-cap: "Difference in Predicted Probabilities in Confidence that voters will be safe to vote in-person between Treatment and Control"

@@

\newpage

>>>>

# custom function to return a certain selection of logistic regression model
# goodness-of-fit statistics
lrm.gof.stats <- function(model, ...){

  out <- data.frame(
    nobs = stats::nobs(model),
    null.deviance = model$null.deviance,
    df.null = model$df.null,
    logLik = as.numeric(stats::logLik(model)),
    deviance = stats::deviance(model),
    df.residual = stats::df.residual(model),
    chisq = as.numeric(model$null.deviance - stats::deviance(model)),
    df = as.numeric(model$df.null - stats::df.residual(model)),
    'P(>chi)' = pchisq(q=model$null.deviance - stats::deviance(model),
                       df = model$df.null - stats::df.residual(model), 
                       lower.tail = F)
  )
  out <- dplyr::as_tibble(out) |>  
  # rename column variables
  dplyr::rename(
    "Num.Obs."      = "nobs",
    "Log.Lik"       = "logLik",
    "Deviance"      = "deviance",
    "Deviance Null" = "null.deviance",
    "DF"            = "df",
    "chisq"          = "chisq",
    "P(>chisq)"     = "P..chi."
  )
  return(out)
}

```{r}
#| label: tinytable-theme

# see here: https://vincentarelbundock.github.io/tinytable/vignettes/theme.html#user-written-themes

theme_mitex <- function(x, ...) {
    fn <- function(table) {
        if (isTRUE(table@output == "typst")) {
          table@table_string <- gsub(
            "\\$(.*?)\\$",
            "#mitex(`\\1`)",
            table@table_string)
        }
        return(table)
    }
    x <- style_tt(x, finalize = fn)
    return(x)
}

@@

Regardless of the measures taken by election officials to boost public confidence in the *trustworthiness* of election administration (e.g., conducting audits, testing election machines), public *trust* and confidence in elections more generally is apt to shift dramatically post-election based on factors such as partisanship, elite rhetoric, particular state policies, and more [@nadeau1993; @carter2024; @coll2024a]. Comparable research finds that confidence in elections is increased through elite rhetoric from fellow partisans affirming legitimacy of elections [@clayton2023], or through the provision of information about the electoral process or other official messaging of election officials themselves [@gaudette2025]. Moreover, prior research has found that evaluation of election workers themselves are an important factor when it comes to levels of public confidence in the electoral process [@hall2007; @hall2009; @claassen2008]. Such studies focused on the quality of the voter experience with reference to the interaction between voter and election worker.

Beyond the general competence of election workers, however, the quality of the voter experience may be influenced merely by prior impressions about *who* comprises election staff and volunteers. Political and other social science researchers have recognized for some time the power that group identity can have over attitudes and perception [@xiao2016; @xiao2012; @vanbavel2021]. As such, we can expect that information identifying the particular groups being recruited to serve as election staff will be enough to improve confidence in elections, and lessen expectations of electoral fraud. The next section elaborates on the proposed theory and why military veterans are of particular interest in this regard.

## Theory: Why Veterans

The theory tested in this study is this: when the public is informed that election officials are actively recruiting veteran military service members to work at election sites as staff or volunteers, their overall confidence in the fairness, accuracy, and integrity of the electoral process will improve. Said another way, insecurities over the integrity of elections administration and safety of voters lessen as the public learns that military service veterans are being sought out by election officials to staff or volunteer at election sites. The stimulating influence of veterans as a group in the abstract is expected to counteract distrust and insecurity pertaining to the institution of elections administration, as well as inspire greater trust in the same. This prompts an obvious question: why would information about efforts to recruit veterans to work as election staff and volunteers have any influence on confidence in elections?

Election officials don't focus on recruiting people of a particular demographic, status, class, or prestige in order to shore up public trust in elections, and they don't discriminate recruitment efforts based on who the public thinks is best suited to secure elections from fraud. Election officials are likely to be agnostic as to who dedicates their time to civil service such as election work. Especially since staffing has been an issue since at least the 1990s [@maidenberg1996; @ferrer2024]. In 2020, such efforts were made far more difficult by the COVID-19 pandemic [@mena2020; @abbate2020a]. There's no special reason to target veterans for recruitment above other groups. Indeed, there's no reason to discriminate recruitment efforts at all if the point is purely to fill staffing vacancies.

There was a sudden push to target veterans for recruitment efforts that arose shortly after the events on Capitol Hill on January 6th, 2021. After the 2020 election, large efforts were made to recruit military veterans and their families to work or volunteer as election staff [@wetheveterans2022; @nflfootballoperations2022]. Prior to that point, young people were sometimes given special mention as targets of election worker recruitment efforts [@powerthepolls2020; @herndon2020]. Generally, however, recruitment efforts cast a wide net, indiscriminate of who applies [@ross2020; @conde2020].

One can speculate that the motivation to associate military veterans with civic engagement and democracy may be intended to counter negative perceptions and impressions given by the proportion of veteran service members arrested for taking part in the events on January 6th [@milton2021; @loewenson2023; @jensen2022a]. Especially with regard to research demonstrating that willingness to support violent efforts to overturn election results (in support of Trump) is, on average, more common among veterans than among matched samples of non-veterans [@pape2024]. This is in addition to a strengthened association portrayed in media outlets between military veterans and militias [@steinhauer2020]. Prior research has substantiated such a connection between veterans and militia groups. @cooter2024 notes from her 3-year ethnographic fieldwork among Michigan militia members that, "...approximately 40% of militia leaders and 30% of members had previous military experience. Most of these veterans actively sought out such groups, as opposed to being recruited by them" [see also @cooter2013]. Thus, countering such associations by promoting a different image of veterans to the mass public and veterans alike seems like a reasonable motivation. Yet such speculation is just that.

The claim that recruiting veterans to work as election staff and volunteers would improve public confidence in elections emerges as non-profit organizations---mainly comprised of veteran and military family members themselves---operate under, and promote, that assumption [@looker2024; @lawrence2024; @wentling2024]. Non-profit groups of late have made incredible strides in targeted recruitment efforts under the notion that veterans are uniquely capable, committed, and best fit to serve the needs of the United States as civil servants and volunteers at election sites. Indeed, this study can be seen as a test of a particular assumption that veterans have about themselves---the assumption that their mere presence and civic engagement will inspire confidence in elections among the public.

Veterans are, after all, civilians. Civilians with shared (but varied) experience and status of prior military service, but who are no more or less qualified for this civil service^[Research is limited and mixed on whether veterans in general are more inclined, however [@nesbit2011].] over non-veteran civilians. To be frank, announcements about concerted efforts to recruit veterans as election workers should not reasonably influence one's trust that votes will be counted more accurately, or that election fraud will not occur; there is no known evidence demonstrating that actually employing veterans as election staff or volunteers further improves the safety, security, and overall integrity of elections in the United States. Moreover, there's no known evidence of positive influence on public perceptions of said outcomes when voters encounter military veterans working as staff or volunteers at elections sites. So, why veterans?

Two considerations underpin the notion that information about veterans will improve trust and quell distrust in elections. First, positive public regard for veterans *in general* is sufficient to elicit greater trust and lower distrust in elections despite the fact that information about veterans as election workers, or even merely information on efforts to recruit veterans, has no bearing on the actual integrity of the electoral process.

An underlying assumption here is that there *is* something special about veterans in the eyes of the public generally, but this doesn't include a proposition on exactly *what* that special something might be in theory. This is partly due to there being limited research on public opinion about former and active military service members distinguished separately from public opinion about the military as an institution in general. Support for the military is generally pretty high, but attitudes about the military in general can't be construed as public opinion about veterans as a group. The evidence that can be gathered concerning public regard for military veterans shows that, generally, veterans are perceived favorably among the general public and all groups express exaggerated support for veterans in more than one way [@kleykamp2018], suggesting that public regard for veterans is prescribed by widespread social norm. Although military recruitment shows a downward trend as of late, public perceptions of veterans are overwhelmingly positive [@kleykamp2023].

Beyond such findings, it is overtly clear that former and active military service members are highly revered nationwide. Veteran's day is a national holiday, and custom prescribes thanking military members for their service. Calls to "support the troops" are distinct from calls to support the politics the troops are deployed to pursue. In 2016, former NFL player Colin Kaepernick went from sitting to kneeling in protest during the playing of the National Anthem upon criticism that sitting in protest implied disrespect for active and former military service members [@haerens2018].

Generally high public regard for veterans is taken together with literature suggesting that public trust in elections is malleable to factors that range in logical relevance to the actual structure or administration of elections. On the one hand, public trust is, indeed, influenced by information specifically about the processes of conducting elections [@gaudette2025]. On the other hand, the public's trust in elections is not beholden to the demonstrated *trustworthiness* of the institution of election administration. Whether generally high public reverence for veterans is due to injunctive social norms or not, what may be posited is simply that positive regard for veterans in general may function sufficiently as a cue to elicit greater trust in elections despite relevance.

Second, veterans are a particularly potent group where mention of one's veteran status seems to have a calming effect on attitudes. More so than just the general positive favor afforded to veterans, the group appears to have a somewhat pacifying effect on negative or otherwise contemptible discrimination, prejudice, and social stigma. For instance, recent research shows that, during his campaign in the 2020 Democratic Primaries, Pete Buttigieg's military background mitigated discrimination against him when he was presented as a veteran married to a man [@magni2024]. Similar research has also found that a candidate's veteran status affords them better evaluations regarding competency in particular issue areas (e.g., war competence) [@teigen2013; @hardy2019]. Moreover, veteran status seems to mitigate or nullify usual stigmas associated with mental illness. That is to say, there is negative stigma associated with mental illness [@corrigan2002] and such stigma incurs labor market discrimination [@hipes2016], but evidence suggests that veteran status overrides such stigma and discrimination [@maclean2014]. Or, in another light, mental illness is seemingly more *understandable* (i.e., permissible) for veterans given the presumptive reasons for their mental strife. And media framing as such plays an important role on public perception [@kleykamp2015]. Relatedly, attitudes about homelessness and PTSD---ranging from perceived causes of homelessness, whether the federal government should dedicate more resources, perceived effectiveness of policies, to compassion for homeless individuals---are generally more favorable for the homeless population among veterans than for the homeless population of general adults [@tsai2021].

Such effects primarily pertain to attitudes about individual people or groups (e.g., a political candidate, prospective employees, the homeless population). The capacity of one's veteran status to pacify prejudice, discrimination, and stigma likely extends out to veterans as a group in a similar fashion. Information that election sites will be staffed by veterans may be enough to ease insecurities about the competence and commitment of the institution to conduct fair and accurate elections.

## Trust and Distrust in Elections

This theory is constrained to public confidence in the administration of elections in the United States. To be clear, the scope conditions in which this theory applies concern public trust and distrust in election administration in the anticipatory period prior to the election event. The theoretical relationship between trust and distrust is important to understand, especially when it comes to the dependent outcome of interest. Before going further, it is necessary to explicate the conception *confidence in elections* with respect to how I've defined trust and distrust in elections.

Trust is, after all, merely *belief* held in the face of uncertainty. Trust is reliance on the veracity of belief in the face of uncertainty. Said another way, trust is *assured belief* dependent on its supposed veracity; a state of certainty with respect to some belief or judgement. It is "*assured*" because it is one's belief that relies on something rather than nothing. Hence quantitative methods of the sciences are replete with things such as *confidence* intervals around our estimates. Therefore, I understand trust to refer to the state or quality of certainty around one's beliefs or expectations; the extent to which one is assured of their expectations regarding what is or what will be the case.

The kind of *trust in elections* I refer to concerns the expectations drawn up from one's judgement about the functioning of the electoral process---the extent to which one is assured in the belief that the electoral process will be fairly and accurately administered. I follow the notion of public trust in elections explicated by @stewart2022,

> "Public trust in U.S. election institutions comes down to whether voting machines accurately record votes, voter registration systems accurately record those eligible to vote, geographic information systems accurately assign voters to voting districts, election-night reporting systems accurately aggregate and communicate election results to officials and the public, and postelection audit and canvassing procedures proceed impartially and in accordance with the law." [-@stewart2022, 236]

That being said, I also account for expectations of electoral fraud, which I refer to as *distrust* in elections distinguished from trust. Distrust in this sense is also a belief, an expectation, about the anticipated functioning of the process. Note, however, that the prefix *dis-* in distrust implies "apart", "lack of, not" or "opposite". Conventionally, distrust is defined simply as the absence of trust, thus denoting a simple inverse and mutually exclusive relationship. However, as it concerns the psychology of an individual, *a lack of trust doesn't necessarily imply opposite expectations or opposing beliefs*. In particular, a lack of trust that votes will be accurately recorded, etc., doesn't necessarily imply inversely proportional expectations that electoral fraud will occur. In other words, although the relationship between trust and distrust is inverse, this relationship is not mutually exclusive.

A person who is certain that electoral fraud will occur clearly lacks trust in elections; an expectation that electoral fraud will occur necessarily implies a lack of trust in elections regardless of positive expressions of trust. However, the reverse isn't also true; expressing a lack of trust in elections *doesn't necessarily imply* an expectation that electoral fraud will occur. Rather, a lack of trust implies an absence of assurance in the belief that the process will be conducted fairly and accurately---a lacking quality of certainty around said belief. An absence of such assurance denotes a state of *insecurity*. Other sources of uncertainty may challenge a person's confidence regarding the integrity of elections independent from anything that would lead them to believe electoral fraud will occur or has occurred. However my use of distrust refers to the assured belief that the outcome will be invalid---that outcomes will be independent of the process as prescribed, thus rendering outcomes as arbitrary. Specifically, distrust implies assurance in the belief that electoral fraud will occur.

As such, trust and distrust represent opposing degrees of confidence along the same spectrum. However, despite the inherent contrast, one can hold both positive degrees of trust and distrust, feeling relatively confident in both respects. Accordingly, *equivalent degrees of trust and distrust cancel out*, rendering one relatively more insecure about their expectations of the future despite the fact that they may report feeling confident in either direction respectively. When trust and distrust are considered reflective of one's confidence about future expectations, then both can be placed along the same spectrum. Therefore, a lack of confidence denotes insecurity.

This means that a person's baseline level of confidence in elections must take into account their present degree of distrust---i.e., their expectations of electoral fraud. Something may cause a person to have greater expectations that election fraud will occur, which in turn will always lower their trust in elections. In contrast, a person's trust in the integrity of the electoral process is reduced by the extent to which they expect that electoral fraud will occur.

Thus, it follows that a person distrusts the integrity of the electoral process to the extent that they are to some extent certain that electoral fraud will occur or has occurred. They are, perhaps, poised to discount election results as invalid or illegitimate in advance. Either way, both trust and distrust refer to some positive degree of confidence (i.e., assurance) regarding one's expectations. They are distinguished by reference to either the normative idea about what *should* come about, or what is *supposed* to come about given the logical structure and integrity of the process. If the structure and integrity of some process is coherent and secure, respectively, then the outcome that follows will be valid. Should the integrity of that process be damaged or opaque, then the process becomes vulnerable and validity of the results are open to challenge. With respect to the administration of U.S. elections, the confident voter expects that votes will be counted as voters intend, election staff will competently administer elections, election technology will be secure from nefarious tampering, and that the process will be fair for all involved.

## Hypotheses

The outcome of interest is *confidence in elections*, which refers to trust in elections---i.e., expectations that elections will be administered fairly and accurately---compensated by distrust in elections---i.e., expectations that electoral fraud will occur. Considering the generally positive perception afforded to veterans by the public, it is reasonable to suspect that efforts that promote recruitment of veteran military service members to work as election staff and volunteers would boost public *confidence in elections administration*. Indeed, that is the primary hypothesis of this study. Although citizens may vote in various ways across the country (e.g., by mail, ballot drop box, in-person), the simple announcement that election officials are engaging in efforts to recruit veterans to work as election staff may boost confidence in elections administration regardless of how, or whether, an individual plans to vote. Formally,

@@

A manipulation check is intended to ensure that the experimental treatment (or manipulation) is received by respondents as the researcher intends. Since this experiment required respondents to read a faux news article, inattention or speeding through vignette presentation would undermine results. To mitigate this as much as possible, the survey included two items as an attention check prior to presenting either the treatment or control vignette. Respondents who failed to select a certain response twice in a row were removed from the survey. Upon presentation of either the treatment or control vignette, survey respondents were unable to progress further in the survey until at least 15 seconds had elapsed.

@@

```{r}
#| label: fig-coef1
#| fig-height: 3.5
#| fig-pos: 'H'
#| fig-cap: "Confidence in Elections by Experiment Condition"

@@

conf.coef1

@fig-coef1 displays standard difference estimates of two models for confidence in elections in Maricopa County, AZ, and one's local area by treatment condition (control condition as reference).

```{r}
#| label: fig-coef2
#| fig-height: 3.5
#| fig-pos: 'H'
#| fig-cap: "Trust and Distrust by Treatment Condition"

@@

conf.coef2

@@

@fig-coef2 displays the same except *confidence* is decomposed into its two components, *trust* and *distrust*. @fig-coef2 merely illustrates the effect of the treatment with respect to the relationship between trust and distrust; ultimately confidence improves.

@@

```{r}
#| label: tbl-legit
#| tbl-cap: "Confidence in Elections by Treatment and Legitimacy Beliefs"
#| tbl-pos: t

@@

The first model in @tbl-legit shows that overall confidence in elections among those who read the treatment vignette was significantly (statistically) greater than those who read the control, on average and netting out the effects of partisanship and legitimacy beliefs. Democratic partisanship remains marginally significant at $90\%$ confidence level ($p = 0.055$) compared to the reference group (Independents). Not believing the 2020 election of Joe Biden was legitimate is negatively associated with confidence in elections. In order to assess the magnitude of this interaction effect of the treatment on those who refuted the 2020 election results, I ran a model where confidence in elections was regressed on the interaction between the treatment and legitimacy beliefs, controlling for partisanship.

@@

While confidence in elections is one concern, it is not the only concern that arises in anticipation of US elections; the public must feel confident that voters are safe to cast a ballot in person. The survey included questions^[See @tbl-6 in Appendix A] that allowed me to assess whether those in the treatment condition expressed less concern for violence and more confidence in voter safety compared to those in the control condition.

```{r}
#| label: barplot-setup
#| include: false

@@

```{r}
#| label: fig-safety1
#| layout-nrow: 2
#| fig-cap: "Safety Concerns by Treatment Condition"
#| fig-subcap:  
#|   - "Concerns for Violence"
#|   - "Confidence in Voter Safety"

@@

Briefly reviewing @fig-safety1, those in the treatment condition felt less concern over the potential for violence compared to those in the control.

As for confidence in voter safety at election sites, it appears that more respondents in the treatment condition expressed that they were "Very confident" compared to those in the control, by approximately $8$ percentage points.

@@

@@

```{r}
#| label: tbl-preds-diff1
#| tbl-colwidths: [80]
#| tbl-pos: 'H'
#| tbl-cap: "Difference in Predicted Probabilities in Concern for violence, Threats of Violence, or Intimidation between Treatment and Control"

@@

Shown in @tbl-preds-diff1, those in the treatment condition had lower probability to express that they were 'Very' or 'Somewhat' concerned for violence in Maricopa County, AZ Elections, yet higher probability of expressing 'Not too' or 'Not at all' concerned.

```{r}
#| label: tbl-preds-diff2
#| tbl-colwidths: [80]
#| tbl-pos: 'H'
#| tbl-cap: "Difference in Predicted Probabilities in Confidence that voters will be safe to vote in-person between Treatment and Control"

@@

As for confidence in voter safety in @tbl-preds-diff2, moving from treatment to control, the difference in predicted probability of a respondent selecting "Very Confident" was $0.08$. Both tables show the significant differences in predicted probabilities at 95$\%$ confidence^[The same models were run for items pertaining to safety concerns in one's local area but results of the treatment were not significantly different from the control. See Appendix E for full results table of ordinal logistic regression]. These findings support both H~4a~ and H~4b~ respectively, permitting rejection of the null for each hypothesis.

## Discussion

The research question emerged in light of the recent surge of efforts nationwide to recruit veterans to serve as election workers [@lawrence2022; @wentling2024] midst an environment where election insecurities were heightened and intensified. Heightened insecurities in anticipation of upcoming elections do not merely pertain to the election outcome (e.g., who wins), but to one's general sense of confidence in the administration of elections, the integrity of the electoral process, and to the safety of casting a ballot in-person. This study demonstrated that information neutral to the trustworthiness of election administration can not only improve trust in elections, but simultaneously reduce distrust in elections. Those who were presented information about veteran recruitment efforts reported greater trust that votes will be counted accurately, that election staff will competently administer elections, that election technology will be secure from nefarious tampering, and that the process will be fair for all involved. Moreover, those presented with information about veteran recruitment efforts reported lower expectations that electoral fraud will occur. They also felt that election sites would be safe for voters to cast a ballot in person and were less concerned about potential violence compared to those who were not presented information about veteran recruitment efforts.

Confidence in elections was greater among those who read the treatment vignette announcing efforts to recruit veterans to work as election staff and volunteers, but this effect was conditional on whether survey questions asked about elections Maricopa County, AZ. When survey questions asked about elections in one's local area, confidence in elections among the treatment group was not significantly different from the control group. However, the sample's confidence in their local area elections was high in general relative to elections in Maricopa County, AZ (@fig-dist). The results presented here suggest that the effect of the veteran cue in the treatment vignette is compensatory. That is, the gap in confidence in local elections and elections in Maricopa County is reduced by the veterans cue, therefore compensating for the otherwise lower confidence and greater insecurity in Maricopa County elections.

A question lingers about whether some antecedent mechanism transmits (i.e., mediates) the effects of the veteran treatment vignette to the outcome. To put simply, this study reiterates that there *is* something special about veterans in particular that elicits greater confidence in elections, but cannot go further in determining exactly *what* that special something might be. Factors that *moderate* the relationship of the treatment on the outcome of interest are likely plenty, but one can only offer conjecture of a *mediator* driving results.

One potential explanation is that veterans, and all those who serve in the military, potentially function as a national symbol. The group is abstract just as the nation is abstract, and the potency of the veteran group emulates the potency of a national symbol. Recall Benedict Anderson's [-@anderson2006] meditation on the Tomb of the Unknown Soldier, "...saturated with ghostly *national* imaginings" [-@anderson2006, 9, emphasis original]. In various studies, national symbols such as the flag, the national anthem, or even mere appeal to national identity have potent, yet variable, effects on different outcomes [@schatz2007; @butz2007; @kemmelmeier2008; @butz2009; @gangl2016; @kalmoe2016; @levendusky2018]. It is plausible that military service members and veterans function in much the same way---as a symbol of national loyalty, pride, and nationalism for people to praise and revere out of a sense of duty prescribed by social norm and reinforced by national sentiment. As such, it may be the case that the treatment effect on confidence in elections is mediated through an affective response to a national symbol---in this case, veterans. The veteran cue may indeed stimulate national sentiment and subsequently boost confidence in elections.

In results presented here, confidence in elections is associated with greater confidence among those in the treatment group, but neither vignette conveys information about the procedures or practices meant to safeguard election integrity; nothing in the treatment suggests to respondents that election integrity would be improved by recruiting veterans. Arguably, the results presented here speak to the "emotional pathway" mentioned in @gaudette2025 so long as we assume that there's an inherent appeal to national sentiment (e.g., nationalism, patriotism) by use of veterans[^13]. Importantly, if the veterans treatment is mediated by stimulation of affect for nation or national identity, then comparable stimulation of national sentiment unrelated to veterans should produce similar results via this "emotional-pathway". However, appeals to national sentiment by use of sterile national symbols (e.g., banal presentation of the U.S. flag, playing of the national anthem) may not have the same affective influence on confidence in elections.

[^13]: The study by @gaudette2025 included an experiment that examined the contrast between "emotional" versus "informational" appeals on trust in elections and beliefs that election fraud is prevalent. Authors presented one of two video messages created by election administration officials in either Virginia or Maricopa County, AZ; the former was "emotions-based" in that it provided little to no information about the voting process or election administration, whereas the latter provided, "...specific information about the processes used to ensure the integrity of the vote with no appeal to emotion or patriotism" [-@gaudette2025, 13].

In a similar vein, it is plausible that certain political apprehensions are nullified insofar as veterans as a group are generally perceived as 'apolitical'---which itself may be enough to engender favor among the mass public [@krupnikov2022]. Current and former military service members may be viewed by the public generally as 'apolitical', especially in comparison to practically any other social group. Indeed, an 'apolitical' reputation has been long desired by military leaders and top brass [@lagrone2019; @garamone2016; @garamone2016a; @baldor2020] despite evidence to the contrary [@betros2001; @mcnerney2006]. Nevertheless, the putative 'apolitical' nature of military service may generalize to veteran service members, and increase confidence in elections based on the assumption that veterans uphold a commitment to ensure election integrity above their own personal loyalties, partisanship, or politics. Prior research has relied on the assumption that such a "*group-based*" norm persists among military members, which prescribes that the military is, or ought to be, apolitical [@mullinix2022]. Although, the extent to which this perception generalizes widely among the mass public about the military in general, or about current and former service members in particular, has yet to be examined. Given recent events of far more overt politicization of the U.S. military [@baldor2025; @baldor2025a; @baldor2025b; @baldor2025c; @thomas2025], it would be useful for future research to consider whether perceptions about the military generally, and military leaders in particular, generalize to perceptions about current and former military service members. Especially insofar as such attitudes relate to confidence in the electoral process and administration.

It is difficult to compare magnitude of effect with other studies given that the item number, item wording, measurement, and computation of confidence in elections is unique across most individual studies to include this one. For instance, @gaudette2025 doesn't compose multiple items into a single composite measure, but subtracts pre-treatment values from post-treatment values for single-item measures with Likert-style response options. Whereas @clayton2023 develops a scale referred to as "Trust in American electoral process" by taking the mean from three items. The item wording of that scale is markedly similar to the items used to compose the trust scale of this study, but the range of response options differs considerably. Thus, while most studies examining public confidence in elections often include the core question "How confidence are you that your vote \[will be/was\] counted as you intended in the most recent election?", the different ways that researchers measure and compute variable scores on the outcome most often produces results that are, unfortunately, sample dependent. The study of public confidence in elections is in desperate need of a standard method of measurement, or even a screened and validated item bank, and this study may be seen as a baby step to that end.

That being said, this study generally emulates treatment effects of comparable studies in the literature. @gaudette2025 found positive effects of official messaging from election officials on various items measuring trust in elections and expectations of electoral fraud. Similarly, @clayton2023, confined to a sample of Republican participants, found positive influence on perceptions of 2020 election legitimacy and trust in the American electoral process from official messaging by Republican politicians reaffirming the 2020 election results.
